Material Methods

### 2019-09-26 doc/MatMeth.Rmd
forgot the change

Hand measures data set generated by using ImageJ/Fiji (v1.52) [@fiji].

@@

framework for analyzing the MLH1 patterns
1. write equation

$X \sim mouse average MLH1 ~ sex + subspecies + random(strain) + \xi $
 can't get the equation in the right format. I think this breaks the pdf kniting.

mixed model, observations of mouse means of mlh1, all data with both male and female observations

subspecies, sex and subsp*sex are fixed effects
strain as random effect (sample of a genetic background from subspecies)
with domesticus male set as the default

strain and sex interaction effect was included to ask / test if the effect of sex varies by strain(genetic background)

this model allows us to estimate and test the fixed effects (sex and subsp)

the random effect of strain in this model can be interpreted as the average standard deviation due to the random strain effect
and sex  strain,

@@

I may have to do more research/reading on testing the random effects.

-specific statistical tests; (means, non-parametric kurts wallis t-test?) (correlation coefficient (function))
linear models lm() from X package.

-specific R packages
-non-parametric  (tests and measures for dealing with count data)
-hypothesis tested (size, DSB number)

>>>>

Measures of meiotic chromosomes generated by using ImageJ/Fiji (v1.52) [@fiji].

@@

We choose to use a **mixed model framework** to interpert the patterns of variation. For interperting patterns of mean MLH1 count within and between house mouse subspecies.

1. write equation

$X \sim mouse average MLH1 ~ sex * subspecies + random(strain) + \xi $

In this model
mixed model, observations of mouse means of mlh1, all data with both male and female observations

The main effects and the interaction of subspecies and sex were coded as fixed.

Subspecies, sex and subsp*sex are fixed effects

Strain was coded as a random effect, since in this experimental design we consider it as a sample of natural genetic variation for each subspecies. We also included a term to allow for strains to have a specific sex effect.

This model allows us to estimate and test the fixed effects (sex and subsp), and variation in mean MLH1 counts due to strain effect.

We use the subspecies term as a way to quantify between group variation (or divergence) and the random strain effect to quantify within group variation (or polymorphism). (in a sex specific manner)

@@

I may have to do more research/reading on testing the random effects.
-specific statistical tests; (means, non-parametric kurts wallis t-test?) (correlation coefficient (function))
linear models lm() from X package.
-specific R packages
-non-parametric  (tests and measures for dealing with count data)
-hypothesis tested (size, DSB number)

### 2019-09-26 doc/MatMeth.tex
forgot the change

---
title: "Material Methods"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_depth: '5'
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '5'
bibliography: refs_MatMeth.bib
---

```{r set.up, echo=FALSE}
#set wd
#load datafile

#common bugs
#html won't compile -- one of my citations in bib was breaking things
```

# Mouse Husbandry

To access natural genetic variation for *Mus musculus*, wild derived inbred strains were used. The Mus musculus PWD/PhJ, PERC/Eij, WSB/EiJ, LEWE/EiJ, MSM/MsJ, MOLF/EiJ, CAST/EiJ, CZECHII/EiJ were purchased from Jackson labs (Maine, USA). The strains of KAZ/TUA, TOM/TUA, AST/TUA, and HMI/TUA were cryo-derived from Biological Resource Center (BRC) at Riken (Ibaraki, Japan).  (https://en.brc.riken.jp).
The related species of *Mus caroli* CAROLI/EiJ and *Mus spretus* SPRET/EiJ were purchased from Jax labs (https://www.jax.org) and *Mus spicilegus* SPIC/Eij from Riken.

 All mice were housed UW-Madison Biotech and MSC facilities - following the protocols.  A breeding colony of wild derived 
*Mus musculus* mice sampled from Gough Island (GI)is maintained at UW Veterinary school facilities.
 Mice were feed on dry standard breeder chow. Some strains sunflower seeds, nestlets and larger cages were added to improve fertility and litter survival. Adult mice were euthanized by CO asphyxiation. Neonate and embryonic mice were euthanized by decapitation following the protocols approved by the Institutional Animal Care and Use Committee at the University of Wisconsin-Madison.

@@

# Tissue Collection and Immunohistochemistry

Spermatocyte spreads were collected and perpared as described in [@peterson_2019] . The majority of mice used were between 5 and 12 weeks, (supplemental table). However some strains had problems breeding, so testes from older mice were collected. We found a (small/no) age effect on MLH1 counts for male mice (Supplemental Figure)).

@@

Meiocyte spreads were made following [@peters_1997]. For testes the tunica was removed and whole testes was incubated in 3ml of hypotonic solution for 45min. For ovaries after dissection, the pair of ovaries were decapsulated in cold PBS and both were incubated in 300ul of hypotonic solution for 45 min.

After incubation, gonads were transferred to sucrose solution -for mastication to make a cell slurry which was transferred to 80ul of 2% PFA solution. Cells were fixed in this solution and dried in a humid chamber at room temperature overnight. The next morning, slides were treated with a photoflow (Kodak, cite) wash.

@@

Explain list recipes for DMC1 since these haven't been listed before

# Image Processing

@@

Repeatability was assessed by X (comparing mulitple counted images)

Measures of meiotic chromosomes generated by using ImageJ/Fiji (v1.52) [@fiji].

total Sc was measured [@wang_2019_sc]

Bivalent level measure were done [@peterson_2019]

# Statisical Analysis

Statistical analysis was performed in R (v3.5.2)[@Rstudio].

@@

We choose to use a **mixed model framework** to interpert the patterns of variation. For interperting patterns of mean MLH1 count within and between house mouse subspecies.

1. write equation

$X \sim mouse average MLH1 ~ sex * subspecies + random(strain) + \xi $

In this model
mixed model, observations of mouse means of mlh1, all data with both male and female observations

The main effects and the interaction of subspecies and sex were coded as fixed.

Subspecies, sex and subsp*sex are fixed effects

Strain was coded as a random effect, since in this experimental design we consider it as a sample of natural genetic variation for each subspecies. We also included a term to allow for strains to have a specific sex effect.

This model allows us to estimate and test the fixed effects (sex and subsp), and variation in mean MLH1 counts due to strain effect.

We use the subspecies term as a way to quantify between group variation (or divergence) and the random strain effect to quantify within group variation (or polymorphism). (in a sex specific manner)

@@

I may have to do more research/reading on testing the random effects.
-specific statistical tests; (means, non-parametric kurts wallis t-test?) (correlation coefficient (function))
linear models lm() from X package.
-specific R packages
-non-parametric  (tests and measures for dealing with count data)
-hypothesis tested (size, DSB number)

Comparison/analysis of variation in MLH1 counts within and between groups was done within a general linear/ mix model frame work, using X package (cite).
With X formula Co count ~ blah + blah, with random strain specific sex effect.

(we used this frame work as proxy for 'polymorphism' and 'divergence' in MLH1 counts across house mouse and closely related species.)

# References

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Material Methods},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Material Methods}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{2019-09-18}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{5}
\tableofcontents
}
\section{Mouse Husbandry}\label{mouse-husbandry}

To access natural genetic variation for \emph{Mus musculus}, wild
derived inbred strains were used. The Mus musculus PWD/PhJ, PERC/Eij,
WSB/EiJ, LEWE/EiJ, MSM/MsJ, MOLF/EiJ, CAST/EiJ, CZECHII/EiJ were
purchased from Jackson labs (Maine, USA). The strains of KAZ/TUA,
TOM/TUA, AST/TUA, and HMI/TUA were cryo-derived from Biological Resource
Center (BRC) at Riken (Ibaraki, Japan). (\url{https://en.brc.riken.jp}).
The related species of \emph{Mus caroli} CAROLI/EiJ and \emph{Mus
spretus} SPRET/EiJ were purchased from Jax labs
(\url{https://www.jax.org}) and \emph{Mus spicilegus} SPIC/Eij from
Riken.

All mice were housed UW-Madison Biotech and MSC facilities - following
the protocols. A breeding colony of wild derived \emph{Mus musculus}
mice sampled from Gough Island (GI)is maintained at UW Veterinary school
facilities. Mice were feed on dry standard breeder chow. Some strains
sunflower seeds, nestlets and larger cages were added to improve
fertility and litter survival. Adult mice were euthanized by CO
asphyxiation. Neonate and embryonic mice were euthanized by decapitation
following the protocols approved by the Institutional Animal Care and
Use Committee at the University of Wisconsin-Madison.

@@

\section{Tissue Collection and
Immunohistochemistry}\label{tissue-collection-and-immunohistochemistry}

Spermatocyte spreads were collected and perpared as described in
(Peterson, Miller, and Payseur 2019) . The majority of mice used were
between 5 and 12 weeks, (supplemental table). However some strains had
problems breeding, so testes from older mice were collected. We found a
(small/no) age effect on MLH1 counts for male mice (Supplemental
Figure)).

@@

Meiocyte spreads were made following (Peters et al. 1997). For testes
the tunica was removed and whole testes was incubated in 3ml of
hypotonic solution for 45min. For ovaries after dissection, the pair of
ovaries were decapsulated in cold PBS and both were incubated in 300ul
of hypotonic solution for 45 min.

After incubation, gonads were transferred to sucrose solution -for
mastication to make a cell slurry which was transferred to 80ul of 2\%
PFA solution. Cells were fixed in this solution and dried in a humid
chamber at room temperature overnight. The next morning, slides were
treated with a photoflow (Kodak, cite) wash.

@@

\section{Image Processing}\label{image-processing}

@@

Hand measures data set generated by using ImageJ/Fiji (v1.52)
(Schindelin et al. 2012).

total Sc was measured (Wang et al. 2019)

Bivalent level measure were done (Peterson, Miller, and Payseur 2019)

\section{Statisical Analysis}\label{statisical-analysis}

Statistical analysis was performed in R (v3.5.2)(Team 2015).

@@

framework for analyzing the MLH1 patterns 1. write equation

\$X \sim mouse average MLH1 \textasciitilde{} sex + subspecies +
random(strain) + \xi \$

mixed model, observations of mouse means of mlh1, all data with both
male and female observations

subspecies, sex and subsp*sex are fixed effects strain as random effect
(sample of a genetic background from subspecies) with domesticus male
set as the default

strain and sex interaction effect was included to ask / test if the
effect of sex varies by strain(genetic background)

this model allows us to estimate and test the fixed effects (sex and
subsp)

the random effect of strain in this model can be interpreted as the
average standard deviation due to the random strain effect and sex
strain,

@@

I may have to do more research/reading on testing the random effects.

-specific statistical tests; (means, non-parametric kurts wallis
t-test?) (correlation coefficient (function)) linear models lm() from X
package.

-specific R packages -non-parametric (tests and measures for dealing
with count data) -hypothesis tested (size, DSB number)

Comparison/analysis of variation in MLH1 counts within and between
groups was done within a general linear/ mix model frame work, using X
package (cite). With X formula Co count \textasciitilde{} blah + blah,
with random strain specific sex effect.

(we used this frame work as proxy for `polymorphism' and `divergence' in
MLH1 counts across house mouse and closely related species.)

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-peters_1997}{}
Peters, Antoine HFM, Annemieke W. Plug, Martine J. van Vugt, and Peter
De Boer. 1997. ``SHORT COMMUNICATIONS A Drying-down Technique for the
Spreading of Mammalian Meiocytes from the Male and Female Germline.''
\emph{Chromosome Research} 5 (1): 66--68.

\hypertarget{ref-peterson_2019}{}
Peterson, April L, Nathan D Miller, and Bret A Payseur. 2019.
``Conservation of the Genome-Wide Recombination Rate in White-Footed
Mice.'' \emph{Heredity}. Nature Publishing Group, 1--16.

\hypertarget{ref-fiji}{}
Schindelin, Johannes, Ignacio Arganda-Carreras, Erwin Frise, Verena
Kaynig, Mark Longair, Tobias Pietzsch, Stephan Preibisch, Curtis Rueden,
Stephan Saalfeld, and Benjamin Schmid. 2012. ``Fiji: An Open-Source
Platform for Biological-Image Analysis.'' \emph{Nature Methods} 9 (7):
676.

\hypertarget{ref-Rstudio}{}
Team, RStudio. 2015. ``RStudio: Integrated Development Environment for
R.'' Boston, MA. \url{http://www.rstudio.com}.

\hypertarget{ref-wang_2019_sc}{}
Wang, RJ, BL Dumont, P Jing, and BA Payseur. 2019. ``A First Genetic
Portrait of Synaptonemal Complex Variation.'' \emph{PLoS Genetics} 15
(8): e1008337--e1008337.

\end{document}
